One Model May Not Fit All: Subgroup Detection Using Model-Based Recursive Partitioning

### 2024-05-13 Raschtree_MH.Rnw
Pruning -> Reversing splits
Second submission J of School Psych
Second submission J of School Psych

\subsubsection{Pruning a Rasch tree}

@@

\subsubsection{Interpretation of a Rasch tree}

@@

The result is presented in Figure~\ref{fig:MHtree2}, which reveals that the Rasch tree is much more concise with a lower number of splits. In addition, in each node the number of items classified as A, B, or C is displayed. For instance, we can see that in node 1 one item is classified in category B (moderate DIF), and 4 items are classified in category C (large DIF).

@@

The result is presented in Figure~\ref{fig:MHtree4}. For node 4 (split in the covariate \texttt{Age}), items 1 (Painting by Andy Warhol) and 9 (Opera not by Mozart: Aida) show DIF and are %(relative to the DIF-free items for this group comparison) 
more difficult to answer for younger respondents ($24 <$ age $\leq 29$), and easier or less difficult to answer for respondents $>29$.

@@

The displayed information again shows that some items show only negligible DIF (category A) in all nodes, while other items show medium (category B) or large (category C) DIF in some or all nodes. The values of the Mantel-Haenszel effect size measure are negative when the item difficulty is lower in the right daughter node than in the left daughter node. For example, as we already observed in Figure~\ref{fig:MHtree3}, items 1, 6, 7, 8, and 9 have medium or large DIF (DIF categories B and C) and are  easier to answer %(relative to the DIF-free items for this group comparison) 
for participants with female or missing gender, which correspond to the right daughter node of node 1. % Item 4 also has a negative sign, indicating that the item difficulty is lower for participants with female or missing gender. However, this item has a smaller DIF effect size. Thus it falls into DIF category A (negligible) and was not relevant for the split in \texttt{Gender}.

>>>>

@@

\subsubsection{Interpreting a Rasch tree}

@@

The result is presented in Figure~\ref{fig:MHtree2}, which reveals that the Rasch tree is much more concise with a lower number of splits. In addition, in each node the number of items classified as A, B, or C is displayed. For instance, we can see that in node 1 one item is classified in category B (moderate DIF), and four items are classified in category C (large DIF).

@@

The result is presented in Figure~\ref{fig:MHtree4}. For node 4 (split in the covariate \texttt{Age}), items 1 (Painting by Andy Warhol) and 9 (Opera not by Mozart: Aida) show DIF and are %(relative to the DIF-free items for this group comparison) 
more difficult to answer for younger male respondents ($24 <$ age $\leq 29$), and easier or less difficult to answer for older male respondents (age $>29$).

@@

The displayed information again shows that some items show only negligible DIF (category A) in all nodes, while other items show medium (category B) or large (category C) DIF in some or all nodes. The values of the Mantel-Haenszel effect size measure are negative when the item difficulty is lower in the right than in the left child node. For example, as we already observed in Figure~\ref{fig:MHtree3}, items 1, 6, 7, 8, and 9 have medium or large DIF (DIF categories B and C) and are  easier to answer %(relative to the DIF-free items for this group comparison) 
for participants with female or missing gender (node 7). % Item 4 also has a negative sign, indicating that the item difficulty is lower for participants with female or missing gender. However, this item has a smaller DIF effect size. Thus it falls into DIF category A (negligible) and was not relevant for the split in \texttt{Gender}.
